---
title: "Exploring NY State Park"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r importing data, echo = FALSE, message = FALSE}
library(tidyverse)
library(httr)
library(plotly)
library(leaflet)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, echo = FALSE, message = FALSE}
res=GET('https://developer.nps.gov/api/v1/parks?limit=500&api_key=B9nDpbkbrb3kSOjz6kXSxMJ3d6MSpUvt1QqYdeyn')

data = res %>% content("text") %>% jsonlite::fromJSON() %>% as_tibble()
```

```{r initial setup}
NPS_data = data %>% unnest(data) %>% unnest(activities,names_sep = '_') %>% unnest(entranceFees,names_sep = '_')
public_use_data = read.csv("data/Query Builder for Public Use Statistics (1979 - Last Calendar Year).csv")
```

### Data Cleaning

```{r}
#main dataset
wildlife_relevant_public_use = 
  public_use_data %>% 
  janitor::clean_names() %>%
  mutate(
    recreation_visits = as.numeric(str_replace_all(recreation_visits,",","")),
    concessioner_camping = as.numeric(str_replace_all(concessioner_camping,",","")),
    tent_campers = as.numeric(str_replace_all(tent_campers,",","")),
    rv_campers = as.numeric(str_replace_all(rv_campers,",","")), 
    backcountry = as.numeric(str_replace_all(backcountry,",","")),
    recreation_visits_total = as.numeric(str_replace_all(recreation_visits_total,",","")),
    concessioner_camping_total = as.numeric(str_replace_all(concessioner_camping_total,",","")),
    tent_campers_total = as.numeric(str_replace_all(tent_campers_total,",","")),
    rv_campers_total = as.numeric(str_replace_all(rv_campers_total,",","")), 
    backcountry_total = as.numeric(str_replace_all(backcountry_total,",",""))
  )
  

#supplemental dataset
wildlife_relevant_NPS_data = 
  NPS_data %>% 
  janitor::clean_names() %>% 
  select(park_code, full_name, states, activities_name, entrance_fees_cost, entrance_fees_description,latitude, longitude, designation) %>% 
  mutate(
    park_code = str_to_upper(park_code),
    entrance_fees_cost = as.numeric(entrance_fees_cost),
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude)
  )
```

# Interactive Map of Parks near NY

```{r}
NY_close = 
  wildlife_relevant_NPS_data %>% 
  filter((states %in% c("NJ,NY","NJ","NY", "PA", "NJ,PA", "RI", "MA", "CT")))

map <- leaflet(NY_close) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    color = "green",
    popup = ~paste("Name:", full_name,
                            "<br> Designation:", designation),
    label = ~paste("Name:", full_name,
                            "Designation:", designation)
  ) %>%
  setView(lng = -74.0060, lat = 40.7128, zoom = 7)

# Display the map
map
```
Discover some of the nearby parks that is either in NY state or within driving distance! When you place your cursor on the circles, you will be able to explore the name of the park and its disposition. 

# NY Popular Parks for Recreation based on yearly visits

```{r, message = FALSE}
wildlife_relevant_public_use %>%
  distinct(year, state, park_name, .keep_all = TRUE) %>% 
  filter(state == "NY") %>% 
  mutate(text_label = str_c("Year: ", year, " Recreation Visits: ", recreation_visits_total)) %>% 
  plot_ly(x  = ~year, y = ~recreation_visits_total, color = ~park_name, text = ~text_label, mode = 'lines+markers')  %>% 
  layout(title = 'Popular Parks for Recreation Visits in NY state', xaxis = list(title = 'Year'), yaxis = list(title = 'Number of Recreation Visits'))
  
```
Ever wondered what parks were most popular around you? This plot demonstrates some popular visits for recreation visits that people chose located within NY state. We have identified that the Gateway National Recreation Area has been the most popular. The two National Monument, Castle Clinton and Statue of Liberty seems to be a popular destination as well. The dip in 2012-2013 and 2020 accurately reflects the context during that time period: the economic recession in 2012-2013 and the pandemic in 2020 impacted this drop. To see more interesting patterns in the parks with lower visitation visit, zoom into the smaller range.

# Most Visited Parks in New York State in 2023

```{r, message = FALSE}
wildlife_relevant_public_use %>%
  distinct(year, state, park_name, .keep_all = TRUE) %>% 
  filter(state == "NY", year == 2023) %>% 
  select(park_name, recreation_visits_total) %>% 
  arrange(desc(recreation_visits_total)) %>% 
  knitr::kable()
```
Now, we wanted to see the trendiest park destination in 2023, and determine where to go for our next travels. We've arranged the popular destination see the most visited park in the past year.

# Top 5 Parks Seasonal Visits in NY state

```{r, message = FALSE}
seasonal_camping = 
  wildlife_relevant_public_use %>% 
  filter(state == "NY") %>% 
  select(year, month, park_name, recreation_visits, tent_campers, rv_campers, backcountry) %>% 
  mutate(season = 
           case_when(
             month %in% c(12, 1, 2) ~ "Winter",
             month %in% c(3,4,5) ~ "Spring",
             month %in% c(6,7,8) ~ "Summmer", 
             month %in% c(9,10,11) ~ "Fall")
  ) %>% 
  group_by(year, season, park_name) %>% 
  summarise(
    total_visit = sum(recreation_visits, na.rm = TRUE),
    tent_camping = sum(tent_campers, na.rm = TRUE),
    rv_camping = sum(rv_campers, na.rm = TRUE),
    backcountry_camping = sum(backcountry, na.rm = TRUE)
  ) %>% 
  mutate(
    total_camping = tent_camping + rv_camping + backcountry_camping)

top_parks <- seasonal_camping %>%
  group_by(park_name) %>%
  summarise(total_visits = sum(total_visit)) %>%
  arrange(desc(total_visits)) %>%
  slice_head(n = 5)

top_parks <- seasonal_camping %>%
  group_by(park_name) %>%
  summarise(total_visits = sum(total_visit)) %>%
  arrange(desc(total_visits)) %>%
  slice_head(n = 5)

# Add visit rate to the dataset
filtered_data <- seasonal_camping %>%
  filter(park_name %in% top_parks$park_name) %>%
  group_by(park_name) %>%
  mutate(
    total_visits = sum(total_visit),  # Total visits for each park
    visit_rate = total_visit / total_visits  # Calculate visit rate
  )

# Create the plot using visit rate
filtered_data %>%
  ggplot(aes(x = reorder(park_name, -total_visits), y = visit_rate, fill = season)) + 
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(
    title = "NY State Park Seasonal Visit Rates (Top 5 Parks)",
    x = "Park Name",
    y = "Visit Rate"
  )+
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8)
  )
```
  <br /> We thought about how people may visit different parks based on the season! This plot shows the visit rate — the proportion of visitors each park gets during different seasons. Fire Island NS dominate summer, reflecting their urban appeal, while quieter parks like Vanderbilt Mansion NHS draw steady attention across multiple seasons. It’s a snapshot of how New Yorkers and tourists alike connect with nature year-round.

# NY Seasonal Trends in Camping for Parks in NY

```{r, message = FALSE}
seasonal_camping = 
  wildlife_relevant_public_use %>% 
  filter(state == "NY") %>% 
  select(park_name, year, month, recreation_visits, tent_campers, rv_campers, backcountry) %>% 
  mutate(season = 
           case_when(
             month %in% c(12, 1, 2) ~ "Winter",
             month %in% c(3,4,5) ~ "Spring",
             month %in% c(6,7,8) ~ "Summmer", 
             month %in% c(9,10,11) ~ "Fall")
  ) %>% 
  group_by(year, season) %>% 
  summarise(
    total_visit = sum(recreation_visits, na.rm = TRUE),
    tent_camping = sum(tent_campers, na.rm = TRUE),
    rv_camping = sum(rv_campers, na.rm = TRUE),
    backcountry_camping = sum(backcountry, na.rm = TRUE),
    total_camping = tent_camping + rv_camping + backcountry_camping
  ) %>% 
  mutate(
    camping_visit_rate = total_camping / sum(total_camping),
    camping_visit_rate = ifelse(is.nan(camping_visit_rate), 0, camping_visit_rate)  # Replace NaN with 0
)
# Calculate camping visit rate grouped by park_name and season
seasonal_camping %>%
  ggplot(aes(x = year, y = camping_visit_rate, fill = season)) + 
  geom_bar(stat = "identity", position = "stack") +  # Stacked bar chart for proportions
  labs(
    title = "Camping Visit Rate by Park and Season in NY",
    x = "Year",
    y = "Camping Visit Rate",
    fill = "Season"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate x-axis labels for clarity
    legend.position = "top"  # Move legend to the top for better visualization
  )
```

<br /> Looking at the camping visit rate across parks in NY state, a notable pattern is the sudden spike in the proportion of summer camping in 2006. This shift could reflect broader cultural or economic changes that year, such as increased interest in outdoor recreation or economic factors encouraging local travel. The summer spike coincides with a national trend in 2006, where families increasingly turned to camping as an affordable vacation alternative during economic recession. This historical moment showcases how external events can shape seasonal camping behavior.

# Revenue generated from Vanderbilit Mansion NHS over the years

```{r, message = FALSE}
NY_public_use= 
  wildlife_relevant_public_use %>% 
  filter(state == "NY") %>% 
  select(park_name, unit_code, year, month, recreation_visits, tent_campers, rv_campers, backcountry, concessioner_camping)

NY_NPS = 
  wildlife_relevant_NPS_data %>% 
  filter(states =="NY") %>% 
  select(park_code, full_name, entrance_fees_cost)

joint_df = left_join(NY_public_use, NY_NPS, by = c("unit_code" = "park_code")) %>% 
  distinct(year, month, park_name, .keep_all = TRUE)

joint_df %>% 
  filter(park_name == "Vanderbilt Mansion NHS") %>% 
  mutate(revenue = recreation_visits* entrance_fees_cost) %>% 
  select(year, month, recreation_visits, revenue) %>% 
  group_by(year) %>% 
  summarise(
    annual_revenue = sum(revenue)
  ) %>% 
  ggplot(aes(x = year, y = annual_revenue, color = annual_revenue)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_gradient(low = "blue", high = "red", name = "Revenue") +
  geom_text(aes(label = ifelse(annual_revenue == max(annual_revenue), round(annual_revenue, 0), "")),
            vjust = -1, color = "black") +
  labs(
    title = "Annual Revenue Generated from Visitors of Vanderbilt Mansion NHS",
    x = "Year",
    y = "Revenue ($)"
  ) +
  theme_minimal()
```
<br /> We were first curious about how much revenue the Vanderbilt Mansion, a National Historic Site, generates each year. What I found instead was a consistent decline in annual revenue over the years, hinting at a possible drop in visitor interest. The plot shows occasional peaks, but the overall trend is unmistakably downward. This raises important questions—are visitor preferences shifting? Is the site losing visibility? These findings highlights the need for fresh approaches to reignite public interest in this historic site and ensure its story continues to resonate with future generations.
